---
title: "Interpolation Grid and IDW"
author: "Jill Deines"
date: "Thursday, December 04, 2014"
output: html_document
---

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE)
opts_chunk$set(echo=TRUE)
opts_chunk$set(fig.path='figures/IDW/')
opts_chunk$set(cache.path='cache/IDW/')

setwd("S:/Projects/2013/Higgins_Lake/Higgins_Bath/KrigingProject")
```

**R Packages Needed*

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(gstat)
```

## Make Interpolation Grids
Interpolation methods in `gstat` require a spatialPixelsDataFrame grid to interpolate to. Because we aren't sure which resolution will have feasible processing times, we are making 3 grids with the following resolutions:

* 1 meter 
* 3 meters
* 10 meters

For now, I'm in charge of the 1 meter grid.

### Preparation: create a lake polygon with donut hole
In order to have our interpolation grid match the lake/island shores precisely, a polygon may be the best bet. Here, I create polygons from the shore points. In ArcGIS, I split the shore points into 1 island shore points file (IslandBorderPoints_utm.shp) and 1 lake shore points shapefile ('ShoreBorderPoitns_utm').

http://stackoverflow.com/questions/21977720/r-finding-closest-neighboring-point-and-number-of-neighbors-within-a-given-rad

```{r lakePolys}
library(rgeos)
# load lake shore points
polyDir <- 'S:/Projects/2013/Higgins_Lake/Higgins_Bath/GIS/lakePolys'
shorepoints <- readOGR(polyDir,'ShoreBorderPoints_utm', verbose=F)

# simply building polygon doesn't work because the points are out of order
sp_poly <- SpatialPolygons(list(Polygons(list(Polygon(shorepoints)), ID=1)))
plot(sp_poly)

# perform a Graham scan to reorder points --------------------------------------
# get projected coords and find further south point
shorepoints$utm.x <- coordinates(shorepoints)[,'coords.x1']
shorepoints$utm.y <- coordinates(shorepoints)[,'coords.x2']
shorepoints$Id <- 1:length(shorepoints$Id)
refPoint <- shorepoints[shorepoints$utm.y == min(shorepoints$utm.y),'Id']

# calculate pairwise distances for all points
pairDists <- gDistance(shorepoints, byid=T)

# starting at refPoint, find the next closest point to make a new index
# start at second point so while loop works..
clockOrder <- c(refPoint$Id, 1179)  # output sorting vector 
index <- 1179       # set starting index

# go through points until back at the beginning
while(index != 1180) {
  distances <- pairDists[index,] # get row for this point from distance matrix
  nextpoint <- as.numeric(names(sort(distances)[2])) # get ID for closest point
  while (nextpoint %in% clockOrder) {
    j <- 3
    nextpoint <- as.numeric(names(sort(distances)[j]))
    j <- j + 1
  }
  clockOrder <- c(clockOrder, nextpoint)             # add to sorting vector
  clockOrder[clockOrder == 1180] <- NA
  index <- nextpoint
}

#ch <- chull(coordinates(shorepoints))
#foo <- shorepoints[c(ch, ch[1]),]




```


```{r makeGrid}
# load full point dataset (with shoreline points)
gisDir <- 'S:/Projects/2013/Higgins_Lake/Higgins_Bath/GIS'
points <- readOGR(gisDir,"All_PointsWithShoreline", verbose=F)

grid <- as(spsample(startingHeads, 50000, type = "regular"), "SpatialPixels")
```

