---
title: "Interpolation Grid and IDW"
author: "Jill Deines"
date: "Thursday, December 04, 2014"
output: html_document
---

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE)
opts_chunk$set(echo=TRUE)
opts_chunk$set(fig.path='figures/IDW/')
opts_chunk$set(cache.path='cache/IDW/')
```

**R Packages Needed*

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(gstat)
library(rgeos)
```

## Make Interpolation Grids
Interpolation methods in `gstat` require a spatialPixelsDataFrame grid to interpolate to. Because we aren't sure which resolution will have feasible processing times, we are making 3 grids with the following resolutions:

* 1 meter 
* 3 meters
* 10 meters

For now, I'm in charge of the 1 meter grid.

### Preparation: create a lake polygon with donut hole
In order to have our interpolation grid match the lake/island shores precisely, a polygon may be the best bet. Here, I create polygons from the shore points. In ArcGIS, I split the shore points into 1 island shore points file (IslandBorderPoints_utm.shp) and 1 lake shore points shapefile ('ShoreBorderPoitns_utm'). I turn the points into polygons below, with code that is wee bit hacky so that the points are ordered correctly to make one polygon.


```{r lakePolysOuter}
# load lake shore points
polyDir <- 'S:/Projects/2013/Higgins_Lake/Higgins_Bath/GIS/lakePolys'
shorepoints <- readOGR(polyDir,'ShoreBorderPoints_utm', verbose=F)

# simply building polygon doesn't work because the points are out of order
sp_poly <- SpatialPolygons(list(Polygons(list(Polygon(shorepoints)), ID=1)))
par(mar=c(.5,.5,.5,.5))
plot(sp_poly)

# perform a Graham scan to reorder points --------------------------------------
# get projected coords and find further south point
shorepoints$utm.x <- coordinates(shorepoints)[,'coords.x1']
shorepoints$utm.y <- coordinates(shorepoints)[,'coords.x2']
shorepoints$Id <- 1:length(shorepoints$Id)
refPoint <- shorepoints[shorepoints$utm.y == min(shorepoints$utm.y),'Id']

# calculate pairwise distances for all points
pairDists <- gDistance(shorepoints, byid=T)

# starting at refPoint, find the next closest point to make a new index
# start at second point so while loop works..
clockOrder <- c(refPoint$Id, 1179)  # output sorting vector 
index <- 1179       # set starting index

# go through points until back at the beginning
while(index != 1180) {
  distances <- pairDists[index,] # get row for this point from distance matrix
  nextpoint <- as.numeric(names(sort(distances)[2])) # get ID for closest point
  j <- 3
  while (nextpoint %in% clockOrder) {
    nextpoint <- as.numeric(names(sort(distances)[j]))
    j <- j + 1
  }
  # this cuts out a few corners. Manually override nextpoint
  if (index == 1095) nextpoint <- 1094    
  if (index == 1913) nextpoint <- 1914 
  if (index == 1914) nextpoint <- 1915 
  if (index == 1915) nextpoint <- 1945
  if (index == 1945) nextpoint <- 1916
  if (index == 1916) nextpoint <- 1921
  if (index == 1921) nextpoint <- 1922
  clockOrder <- c(clockOrder, nextpoint)             # add to sorting vector
  clockOrder[clockOrder == 1180] <- NA
  index <- nextpoint
}

# add in 1180
clockOrder <- clockOrder[1:(length(clockOrder)-1)]
clockOrder[is.na(clockOrder)] <- 1180

# subset points for ones in clockOrder, and sort by clockOrder
shore2 <- shorepoints[shorepoints$Id %in% clockOrder,]
shore2 <- shore2[match(clockOrder, shore2$Id),]

# now make a polygon -----------------------------------------------------------
poly <- SpatialPolygons(list(Polygons(list(Polygon(shore2)), ID=1)))
poly <- SpatialPolygonsDataFrame(poly, data=data.frame(ID=1))
proj4string(poly) <- proj4string(shorepoints)
par(mar=c(.5,.5,.5,.5))
plot(shorepoints, col='red', cex=.5)
plot(poly, add=T)

# writeOGR(poly, polyDir,'ShorePoly_noIsland_utm', driver = 'ESRI Shapefile')
```

Awesome, now let's do the island.

```{r lakePolysInner}
# load lake shore points
polyDir <- 'S:/Projects/2013/Higgins_Lake/Higgins_Bath/GIS/lakePolys'
islandpoints <- readOGR(polyDir,'IslandBorderPoints_utm', verbose=F)

# simply building polygon doesn't work because the points are out of order
ipoly <- SpatialPolygons(list(Polygons(list(Polygon(islandpoints)), ID=1)))
par(mar=c(.5,.5,.5,.5))
plot(ipoly)

# perform a Graham scan to reorder points --------------------------------------
# get projected coords and find further south point
islandpoints$utm.x <- coordinates(islandpoints)[,'coords.x1']
islandpoints$utm.y <- coordinates(islandpoints)[,'coords.x2']
islandpoints$Id <- 1:length(islandpoints$Id)
refPoint <- islandpoints[islandpoints$utm.y == min(islandpoints$utm.y),'Id']

# calculate pairwise distances for all points
pairDists <- gDistance(islandpoints, byid=T)

# starting at refPoint, find the next closest point to make a new index
# start at second point so while loop works..
nextpoint1 <- as.numeric(names(sort(pairDists[refPoint$Id,])[2]))
clockOrder <- c(refPoint$Id, nextpoint1)  # output sorting vector 
index <- nextpoint1       # set starting index

# go through points until back at the beginning
while(index != refPoint$Id) {
  distances <- pairDists[index,] # get row for this point from distance matrix
  nextpoint <- as.numeric(names(sort(distances)[2])) # get ID for closest point
  j <- 3
  while (nextpoint %in% clockOrder) {
    nextpoint <- as.numeric(names(sort(distances)[j]))
    j <- j + 1
  }
  # this cuts out a few corners. Manually override nextpoint 
  clockOrder <- c(clockOrder, nextpoint)             # add to sorting vector
  clockOrder[clockOrder == refPoint$Id] <- NA
  index <- nextpoint
}

# add in 1180
clockOrder <- clockOrder[1:(length(clockOrder)-1)]
clockOrder[is.na(clockOrder)] <- refPoint$Id

# subset points for ones in clockOrder, and sort by clockOrder
#shore2 <- shorepoints[shorepoints$Id %in% clockOrder,]
island <- islandpoints[match(clockOrder, islandpoints$Id),]

# now make a polygon -----------------------------------------------------------
polyis <- SpatialPolygons(list(Polygons(list(Polygon(island)), ID=1)))
polyis <- SpatialPolygonsDataFrame(polyis, data=data.frame(ID=1))
proj4string(polyis) <- proj4string(islandpoints)
par(mar=c(.5,.5,.5,.5))
plot(islandpoints, col='red', cex=.5)
plot(polyis, add=T)

# writeOGR(polyis, polyDir,'IslandPoly_utm', driver = 'ESRI Shapefile')
```

```{r makeGrid, eval=FALSE}
# load full point dataset (with shoreline points)
gisDir <- 'S:/Projects/2013/Higgins_Lake/Higgins_Bath/GIS'
points <- readOGR(gisDir,"All_PointsWithShoreline", verbose=F)

grid <- as(spsample(startingHeads, 50000, type = "regular"), "SpatialPixels")
```

